Analysis of Major Crimes in Toronto And Strategies for Safety Amidst Increasing Numbers

### 2024-01-25 outputs/paper/paper.qmd
Final check and correct typo

```{r}
#| include: false
# Explore
# Test
clean_mci_data$report_year |> min() == 2014
clean_mci_data$report_year |> max() == 2023
clean_mci_data$occ_year |> unique()
any(duplicated(clean_mci_data$x_id))

@@

```{r}
#| message: false
#| echo: false
#| label: tbl-clean_mci_data
#| tbl-cap: The Number of Major Crimes occurred and the Month with the Most Crimes (2014-2023)
# Create a table with year, 2014 to 2023, the number of crimes in each year, and the month with the most crimes correspondingly.
clean_mci_data |>
  filter(occ_year >= 2014) |>
  group_by(occ_year, occ_month) |>
  summarise(n = n()) |>
  group_by(occ_year) |>
  summarise(
    total_number_of_crimes = sum(n),
    the_most_frequent_month = occ_month[which.max(n)]
  ) |>
  distinct(occ_year, .keep_all = TRUE) |>
  select(occ_year, total_number_of_crimes, the_most_frequent_month) |> 
  kable(
    col.names = c("Year", "Total Number of Crimes", "Month with the Most Crimes"),
    booktabs = TRUE
  )
```
@tbl-clean_mci_data provides a brief summary of the total number of crimes from 2014 to 2023. Over this period, the overall number of crimes shows a consistent increase, except for the years 2020 and 2021, which have a decline probably due to the impact of the pandemic. However, in 2022 and 2023, there is a significant increase by more than 6000. In each year, the month with the highest number of crimes tends to be either July, August, or October, concentrating primarily in the summer-fall period.

```{r}
#| message: false
#| echo: false
#| label: fig-types_of_crimes
#| fig-cap: The Number of Crimes Reported, Classified by Crime Types (2014-2023)
clean_mci_data |>
  ggplot(mapping = aes(x = factor(report_year), fill = mci_category)) +
  geom_bar(position = "stack", stat = "count") +
  labs(x = "Year", y = "Total Number of Crimes", fill = "Types of Major Crimes") +
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set3")
```
Factors contributing to this uptrend are multifaceted, ranging from societal shifts associated with the pandemic to economic uncertainties. In particular, I will discuss the distribution of crimes from three aspects: types of crime, regions and time in a day.

@@

```{r}
#| include: false
clean_mci_data <- clean_mci_data |>
  mutate(
    division_area =
      case_match(
        division,
        "D43" ~ "Scarborough",
        "D42" ~ "Scarborough",
        "D41" ~ "Scarborough",
        "D53" ~ "Toronto-East York",
        "D55" ~ "Toronto-East York",
        "D52" ~ "Toronto-East York",
        "D51" ~ "Toronto-East York",
        "D54" ~ "Toronto-East York",
        "D32" ~ "North York",
        "D33" ~ "North York",
        "D31" ~ "North York",
        "D22" ~ "Etobicoke",
        "D23" ~ "Etobicoke",
        "D14" ~ "Etobicoke-York",
        "D11" ~ "Etobicoke-York",
        "D12" ~ "Etobicoke-York",
        "D13" ~ "Etobicoke-York",
        "NSA" ~ "Other"))
```

```{r}
#| include: false
clean_mci_data <- clean_mci_data |> 
  mutate(
    weekday_weekend =
      case_match(
        occ_dow,
        "Wednesday" ~ "Weekday",
        "Tuesday" ~ "Weekday",
        "Thursday" ~ "Weekday",
        "Sunday" ~ "Weekend",
        "Monday" ~ "Weekday",
        "Friday" ~ "Weekday",
        "Saturday" ~ "Weekend",
        "" ~ "Other"))
clean_mci_data
```
There are 18 police divisions, which are D11, D12, D13, D14, D22, D23, D31, D32, D33, D41, D42, D43, D51, D52, D53, D54, D55, and NSA. I put them into 6 geographical regions, Etobicoke-York, Etobicoke, North York, Scarborough, Toronto-East York, and other as shown in @tbl-division_area. Based on @fig-area_of_crimes, Toronto-East York reports the highest amount of crimes, and Etobicoke-York has the lowest. North York, Scarborough, and Etobicoke-York have similar numbers, but the rate of increase in North York is higher than that observed in the other two. It is crucial to recognize that, though Toronto-East York records the most crimes, it also has the largest number of residents. Therefore, it could be biased to claim that Toronto-East York has the highest crime rate, or North York is getting dangerous, as North York has witnessed a growing population in recent years. Moreover, when creating this dataset, one occurrence number may have a few rows with different major crime indicators types used to categorize the occurrence. Therefore, some incidents of 'Auto Theft,' 'Robbery,' and 'Theft Over' may be potentially recorded multiple times under 'Break and Enter.' Consequently, the number of crimes classified as just ‘Break and Enter’ is overestimated, the actual number may be lower than what is shown now in the dataset.
```{r}
#| message: false
#| echo: false
#| label: tbl-division_area
#| tbl-cap: Police divisions classified into six areas (including other)

clean_mci_data |>
  group_by(division_area) |>
  summarise(division = toString(unique(division))) |> kable()

@@

```{r}
#| message: false
#| echo: false
#| label: fig-area_of_crimes
#| fig-cap: Total Number of Crimes Occurred in Different Police Division Areas (2014-2023)
clean_mci_data |>
  filter(occ_year >= 2014) |>
  count(occ_year, division_area) |>
  ggplot(aes(x = occ_year, y = n, color = division_area)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Year",
       y = "Total Number of Crimes",
       color = "Division Area")+
  scale_x_continuous(breaks = seq(2014, 2023, by = 1))
```

```{r}
#| message: false
#| echo: false
#| label: fig-time_of_crimes
#| fig-cap: Total Number of Crimes Occurred in a Day
clean_mci_data |>
  filter(weekday_weekend == 'Weekday' | weekday_weekend == 'Weekend') |>
  count(occ_hour, weekday_weekend) |>
  ggplot(mapping = aes(x = occ_hour, y = n, color = weekday_weekend)) +
  geom_line() +
  theme(legend.position = "right") +
  labs(x = "Time In a Day", y = "Number of Crimes", color = "Weekday or Weekend") +
  scale_color_manual(values = c("Weekday" = "steelblue3", "Weekend" = "indianred3"))+
  scale_x_continuous(breaks = seq(0, 23, by = 1))
```

>>>>

```{r}
#| include: false
# Explore
# Tests
clean_mci_data$report_year |> min() == 2014
clean_mci_data$report_year |> max() == 2023
clean_mci_data$occ_year |> unique()
any(duplicated(clean_mci_data$x_id))

@@

```{r}
#| message: false
#| echo: false
#| label: tbl-clean_mci_data
#| tbl-cap: The Number of Major Crimes occurred and the Month with the Most Crimes (2014-2023)
# Create a table with years, 2014 to 2023, the number of crimes in each year, and the month with the most crimes correspondingly.
clean_mci_data |>
  filter(occ_year >= 2014) |>
  group_by(occ_year, occ_month) |>
  summarise(n = n()) |>
  group_by(occ_year) |>
  summarise(
    total_number_of_crimes = sum(n),
    the_most_frequent_month = occ_month[which.max(n)]
  ) |>
  distinct(occ_year, .keep_all = TRUE) |>
  select(occ_year, total_number_of_crimes, the_most_frequent_month) |> 
  kable(
    col.names = c("Year", "Total Number of Crimes", "Month with the Most Crimes"),
    booktabs = TRUE
  )
```
@tbl-clean_mci_data provides a brief summary of the total number of crimes from 2014 to 2023. Over this period, the overall number of crimes shows a consistent increase, except for the years 2020 and 2021, which have a decline probably due to the impact of the pandemic. However, in 2022 and 2023, there is a significant increase by more than 6000. In each year, the month with the highest number of crimes tends to be either July, August, or October, concentrating primarily in the summer-fall period.

```{r}
#| message: false
#| echo: false
#| label: fig-types_of_crimes
#| fig-cap: The Number of Crimes Reported, Classified by Crime Types (2014-2023)
# Create a histogram with numbers of crimes classified by crime types from 2014 to 2023
clean_mci_data |>
  ggplot(mapping = aes(x = factor(report_year), fill = mci_category)) +
  geom_bar(position = "stack", stat = "count") +
  labs(x = "Year", y = "Total Number of Crimes", fill = "Types of Major Crimes") +
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set3")
```
Factors contributing to this uptrend are multifaceted, ranging from societal shifts associated with the pandemic to economic uncertainties. In particular, I will discuss the distribution of crimes from three aspects: types of crime, regions and time in a day.

@@

```{r}
#| include: false
# Create a new variable division_area, which put 18 police divisions (including a NSA) into 6 regions of Toronto (including Other for NSA)
clean_mci_data <- clean_mci_data |>
  mutate(
    division_area =
      case_match(
        division,
        "D43" ~ "Scarborough",
        "D42" ~ "Scarborough",
        "D41" ~ "Scarborough",
        "D53" ~ "Toronto-East York",
        "D55" ~ "Toronto-East York",
        "D52" ~ "Toronto-East York",
        "D51" ~ "Toronto-East York",
        "D54" ~ "Toronto-East York",
        "D32" ~ "North York",
        "D33" ~ "North York",
        "D31" ~ "North York",
        "D22" ~ "Etobicoke",
        "D23" ~ "Etobicoke",
        "D14" ~ "Etobicoke-York",
        "D11" ~ "Etobicoke-York",
        "D12" ~ "Etobicoke-York",
        "D13" ~ "Etobicoke-York",
        "NSA" ~ "Other"))
```

```{r}
#| include: false
# Create a new variable weekend_weekday, which indicates whether the occurrence happen on a weekday or a weekend
clean_mci_data <- clean_mci_data |> 
  mutate(
    weekday_weekend =
      case_match(
        occ_dow,
        "Wednesday" ~ "Weekday",
        "Tuesday" ~ "Weekday",
        "Thursday" ~ "Weekday",
        "Sunday" ~ "Weekend",
        "Monday" ~ "Weekday",
        "Friday" ~ "Weekday",
        "Saturday" ~ "Weekend",
        "" ~ "Other"))
clean_mci_data
```
There are 18 police divisions, which are D11, D12, D13, D14, D22, D23, D31, D32, D33, D41, D42, D43, D51, D52, D53, D54, D55, and NSA. I put them into 6 geographical regions, Etobicoke-York, Etobicoke, North York, Scarborough, Toronto-East York and other, as shown in @tbl-division_area. Based on @fig-area_of_crimes, Toronto-East York reports the highest amount of crimes, and Etobicoke-York has the lowest. North York, Scarborough, and Etobicoke-York have similar numbers, but the rate of increase in North York is higher than that observed in the other two. It is crucial to recognize that, though Toronto-East York records the most crimes, it also has the largest number of residents. Therefore, it could be biased to claim that Toronto-East York has the highest crime rate, or North York is getting dangerous, as North York has witnessed a growing population in recent years. Moreover, when creating this dataset, one occurrence number may have a few rows with different major crime indicators types used to categorize the occurrence. Therefore, some incidents of 'Auto Theft,' 'Robbery,' and 'Theft Over' may be potentially recorded multiple times under 'Break and Enter.' Consequently, the number of crimes classified as just ‘Break and Enter’ is overestimated, the actual number may be lower than what is shown now in the dataset.
```{r}
#| message: false
#| echo: false
#| label: tbl-division_area
#| tbl-cap: Police divisions classified into six areas (including other)
# Create a table to show the realation between the new variable division_area and division
clean_mci_data |>
  group_by(division_area) |>
  summarise(division = toString(unique(division))) |> kable()

@@

```{r}
#| message: false
#| echo: false
#| label: fig-area_of_crimes
#| fig-cap: Total Number of Crimes Occurred in Different Police Division Areas (2014-2023)
# Visualize the number of crimes occurred in different police division areas
clean_mci_data |>
  filter(occ_year >= 2014) |>
  count(occ_year, division_area) |>
  ggplot(aes(x = occ_year, y = n, color = division_area)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Year",
       y = "Total Number of Crimes",
       color = "Division Area")+
  scale_x_continuous(breaks = seq(2014, 2023, by = 1))
```

```{r}
#| message: false
#| echo: false
#| label: fig-time_of_crimes
#| fig-cap: Total Number of Crimes Occurred in a Day
# Create a line chart showing the total number of crimes occurred in 24 hours in a day
clean_mci_data |>
  filter(weekday_weekend == 'Weekday' | weekday_weekend == 'Weekend') |>
  count(occ_hour, weekday_weekend) |>
  ggplot(mapping = aes(x = occ_hour, y = n, color = weekday_weekend)) +
  geom_line() +
  theme(legend.position = "right") +
  labs(x = "Time In a Day", y = "Number of Crimes", color = "Weekday or Weekend") +
  scale_color_manual(values = c("Weekday" = "steelblue3", "Weekend" = "indianred3"))+
  scale_x_continuous(breaks = seq(0, 23, by = 1))
```
